Thesis: draft

### 2021-04-23 draft.tex
latest update

\subsection{Starting conditions $P_0(0) = 0.5$ and $P_1(0)=0.5$}
\label{50-50}
We start with the following opinion distribution: a fraction of nodes with opinion 0 equal to 0.5 ($P_0(0) = 0.5$) and a fraction of nodes with opinion 1 equal to 0.5 ($P_1(0) = 0.5$). The opinions are randomly distributed over the network. The opinion dynamics model discussed in Chapter \ref{chap4}, Section \ref{opinionmodel} % correct section??
is used. We set the activation probability equal to $p_{\text{act}} = 0.1$ and the actual time-line of each user has a maximum length equal to $S=20$.\\
Four different network models are implemented, each with a total number of nodes equal to $N = 10^3$ and an average degree around $\left<k\right> \sim 10$
	\item The Watts-Strogatz model (WS) with mean degree $K=10$; two different values of the rewire probability $\beta$ are used, each of them has a slightly different value of the average clustering coefficient, namely $\beta = 0.01$ with $\left<C\right> \sim 0.60$ and $\beta = 0.06$ with $\left<C\right> \sim 0.55$. This model is used to investigate the effect of clustering.
	\item The stochastic block model (SBM) with 10 communities of size 100; two networks with different modularities $Q$ are investigated: a high modularity network, $Q \sim 0.9$, with an edge probability inside the communities equal to $p_{\text{cl}} = 0.1$ and an edge probability between communities equal to $p_{\text{add}} = 0.001$, and a low modularity network ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q \sim 0.2$). This model is used to investigate the effect of a community structure.
Figure \ref{ev_op_50_50} shows the results for the evolution of the fraction of nodes with opinion 0, $P_0(t)$, over time, together with the standard deviation $\sigma$. Two different filtering algorithms are compared to each other, namely the PR method and the REC method (see Chapter \ref{chap3}, Section \ref{filter} for a detailed explanation). None of the network topologies, nor the filtering algorithms, is able to break the initial $50/50$ opinion distribution. Having a high clustering or a clear community structure has no impact on the prevalence of the opinions. The status quo is maintained and the system is not pushed towards a different equilibrium.
  	\includegraphics[width=8cm, height=7cm]{fraction_of_opinions_ER_001_50-50_PR_REC.png}
  	\includegraphics[width=8cm, height=7cm]{fraction_of_opinions_WS_50-50_PR_REC.png}
    \includegraphics[width=8cm, height=7cm]{fraction_of_opinions_SBM_10x100_01-0001_003-0008_50-50_PR_REC.png}
    \includegraphics[width=8cm, height=7cm]{fraction_of_opinions_SBM-WS_10x100_10-001-0001_50-50_PR_REC.png}
  \caption[Evolution of the fraction of nodes with opinion 0 ($P_0(t)$) over time for an initial $50/50$ opinion distribution.]{Evolution of the fraction of nodes with opinion 0 ($P_0(t)$) over time for an initial $50/50$ distribution. Four different network models are used, each with an average degree around $\left<k\right> \sim 10$ and a total number of nodes $N = 10^3$: \textbf{(a)} Erd\H{o}s-R\'{e}nyi model (ER), $p=0.01$; \textbf{(b)} Watts-Strogatz model (WS), $K = 10$, two slightly different values of the clustering coefficient are used ($\left<C\right> \sim 0.55$ and $\left<C\right> \sim 0.60$); \textbf{(c)} Stochastic block model (SBM) with 10 communities of size 100, a high modularity case ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001,\ Q \sim 0.9$) and a low modularity case ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q \sim 0.2$) are depicted; \textbf{(d)} Stochastic Watts-Strogatz block model (SBM-WS) with 10 communities of size 100, each community has the following WS parameters: $K = 10,\ \beta = 0.01$ and $p_{\text{add}} = 0.001$, $\left<C\right> \sim 0.55$ and $Q \sim 0.9$. For each model two different filtering algorithms are compared to each other: the REC and the PR filtering algorithm. Each plot is the average of $10^2$ independent simulations and data points are shown for every 50 time steps to improve visualization.}%more explanation?
\label{ev_op_50_50}

@@

Figure \ref{echo_50_50} shows the distribution of the fraction of friends (nearest neighbors, nn) of node $i$ with the same opinion 1 at $t=500$, $\left<P_1^{\text{nn}}\right>$. The value shown on the $y$-axis, $F_N(\left<P_1^{\text{nn}}\right>)$, is the result of dividing the average distribution of opinions in each neighborhood at $t=500$ by its corresponding value at $t=0$. The region on the $x$-axis close to zero describes neighborhoods in which none, or very few users, adopt opinion 1 (these are thus neighborhoods in which most users adopt opinion 0). Conversely, the region on the $x$-axis close to one describes neighborhoods in which a majority of the users adopt opinion 1. The distribution is obtained by imposing ten equispaced bins, $\Delta x = 0.1$ \cite{Perra2019}. Error bars are omitted for the sake of an improved visualization.\\
These types of plots are perfectly suited to investigate the formation of opinion clusters (echo chambers) and effects such as polarization in the system under study.\\

@@

@@

  	\includegraphics[width=8cm, height=7cm]{fraction_of_opinions_ER_001_50-50_PR_REC_20-80_8x7.png}

@@

  	\includegraphics[width=8cm, height=7cm]{fraction_of_opinions_WS_10-006_PR_REC_20-80_8x7.png}
    \label{ws_20-80_op}
    \includegraphics[width=8cm, height=7cm]{fraction_of_opinions_SBM_01-0001_003-0008_PR_REC_20-80_8x7.png}
    \includegraphics[width=8cm, height=7cm]{fraction_of_opinions_SBM-WS_10-001-0001_PR_REC_20-80_8x7.png}
    \label{sbm-ws_20-80_op}

@@

  \caption[Evolution of the fraction of nodes with opinion 0 ($P_0(t)$) over time for an initial $20/80$ opinion distribution.]{Evolution of the fraction of nodes with opinion 0 ($P_0(t)$) over time for an initial $20/80$ distribution. Four different network models are used, each with an average degree around $\left<k\right> \sim 10$ and a total number of nodes $N = 10^3$: \textbf{(a)} Erd\H{o}s-R\'{e}nyi model (ER), $p=0.01$; \textbf{(b)} Watts-Strogatz model (WS) with $K = 10$, $\beta = 0.06$ and $\left<C\right> \sim 0.55$; \textbf{(c)} Stochastic block model (SBM) with 10 communities of size 100, a high modularity case ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001,\ Q \sim 0.9$) and a low modularity case ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q \sim 0.2$) are depicted; \textbf{(d)} Stochastic Watts-Strogatz block model (SBM-WS) with 10 communities of size 100, each community has the following WS parameters: $K = 10,\ \beta = 0.01$ and $p_{\text{add}} = 0.001$, $\left<C\right> \sim 0.55$ and $Q \sim 0.9$. For each model two different filtering algorithms are compared to each other: the REC and the PR filtering algorithm. Each plot is the average of $10^2$ independent simulations and data points are shown for every 50 time steps to improve visualization.}
\label{ev_op_20_80}

@@

\begin{figure}[H]
  \begin{subfigure}[b]{0.49\textwidth}
  	\includegraphics[width=8cm, height=7cm]{fraction_of_opinions_WS_1000-6_PR_20-80_8x7.png}
    \caption{WS, $N = 10^3,\ K=6,\ \beta=0.01$}
    \label{ws1000}
  \end{subfigure}
  \begin{subfigure}[b]{0.49\textwidth}
  	\includegraphics[width=8cm, height=7cm]{fraction_of_opinions_WS_10000-6_PR_20-80_8x7.png}
    \caption{WS, $N = 10^4,\ K=6,\ \beta=0.01$}
    \label{ws10000}
  \end{subfigure}
  \captionsetup{format=plain}
  \caption[Evolution of the fraction of nodes with opinion 0 ($P_0(t)$) over time for an initial $20/80$ opinion distribution for a WS model with $N=10^3$ or $N=10^4$, $K=6$ and $\beta = 0.01$, PR.]{Evolution of the fraction of nodes with opinion 0 ($P_0(t)$) over time for an initial $20/80$ distribution. The Watts-Strogatz network is depicted with \textbf{(a)} $N=10^3,\ K=6,\ \beta=0.01$ and \textbf{(b)} $N=10^4,\ K=6,\ \beta=0.01$. In both cases the average clustering coefficient is around $\left<C\right> \sim 0.58$ and the PR filtering method is used. Each plot is the average of $10^2$ independent simulations and data points are shown for every 50 time steps to improve visualization.}
\label{ev_op_WS_k=6}
\end{figure}
\newpage
\noindent
\textbf{Echo chambers}\\
  	\includegraphics[width=8cm, height=7cm]{echo_chambers_no_stubb_ER_001_20-80_PR_REC.png}
    \label{er_echo_20-80}
  	\includegraphics[width=8cm, height=7cm]{echo_chambers_no_stubb_WS_10_006_20-80_PR_REC_8x7.png}
    \label{ws}
    \includegraphics[width=8cm, height=7cm]{echo_chambers_no_stubb_SBM_01-0001_003-0008_10x100_20-80_PR_REC_8x7.png}
  	\label{sbm}    
    \includegraphics[width=8cm, height=7cm]{echo_chambers_no_stubb_SBM-WS_10_001_0001_10x100_20-80_PR_REC_8x7.png}
    \label{sbm-ws}
  \caption[Distribution of the fraction of friends (nearest neighbors, nn) of node $i$ with the same opinion 1 at $t=500$, $\left<P_1^{\text{nn}}\right>$, for an initial $20/80$ opinion distribution.]{Distribution of the fraction of friends (nearest neighbors, nn) of node $i$ with the same opinion 1 at $t = 500$. The distribution is normalized by dividing for the same quantity computed at $t=0$. Four different network models are used, each with a total number of nodes $N = 10^3$ and an average degree around $\left<k\right> \sim 10$: \textbf{(a)} Erd\H{o}s-R\'{e}nyi model (ER), $p=0.01$; \textbf{(b)} Watts-Strogatz model (WS) with $K = 10$, $\beta = 0.06$ and $\left<C\right> \sim 0.55$; \textbf{(c)} Stochastic block model (SBM) with 10 communities of size 100, a high modularity case ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001,\ Q \sim 0.9$) and a low modularity case ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q \sim 0.2$) are depicted; \textbf{(d)} Stochastic Watts-Strogatz block model (SBM-WS) with 10 communities of size 100, each community has the following WS parameters: $K = 10,\ \beta = 0.01$ and $p_{\text{add}} = 0.001$, $\left<C\right> \sim 0.55$ and $Q \sim 0.9$. For each model two different filtering algorithms are compared to each other: the REC and the PR filtering algorithm. The starting conditions are $P_0(0) = 0.2$ and $P_1(0) = 0.8$. Each plot is the average of $10^2$ independent simulations.}
\label{echo_20_80}

@@

Lets first discuss the results for the Erd\H{o}s-R\'{e}nyi model and the stochastic block model. When using the REC method, the formation of echo chambers around the majority opinion 1 is hampered. This is somewhat surprising, since $80\%$ of the nodes carry this opinion and is probably a consequence of the absence of clustering and the presence of short average path lengths in these models. There is even a little spike around $\left<P_1^{\text{nn}}\right> = 0.1$, which indicates a small increase in the fraction of nodes with the majority of their neighbors carrying the minority opinion 0. From figure \ref{ev_op_20_80} it was already concluded that the REC method doesn't break the prevalence of the opinions. The opinion distribution is however slightly shifted from the case where the minority opinion 0 is more or less randomly distributed over the network to the case where the minority opinion 0 groups together in opinion clusters. \\

@@

% talk about small spike around 1 for PR method??
% explain black dotted line??

@@

@@

In Figure \ref{ev_op_50_50_stubb_REC} the evolution of the fraction of nodes with opinion 0 for an initial $50/50$ opinion distribution is shown. Data points are shown together with the standard deviation $\sigma$. This figure shows the results for the REC filtering algorithm. The four network models (ER, WS, SBM (high and low modularity) and SBM-WS) are compared to each other. For each model we look at three different fractions of completely stubborn people: no stubborn people, fracRes $=0$; a fraction of completely stubborn people equal to fracRes $=0.4$ and a fraction of completely stubborn people equal to fracRes $=0.8$. Introducing a fraction of stubborn people is not able to break the initial $50/50$ opinion distribution for neither of the network models.\\

\begin{figure}[H]
  \begin{subfigure}[b]{0.49\textwidth}
  	\includegraphics[width=8cm, height=7cm]{fraction_of_opinions_ER_frac_compl_stubb_REC_8x7.png}
    \caption{ER}
  \end{subfigure}
  \begin{subfigure}[b]{0.49\textwidth}
  	\includegraphics[width=8cm, height=7cm]{fraction_of_opinions_WS_frac_compl_stubb_REC_8x7.png}
    \caption{WS}
  \end{subfigure}
  \begin{subfigure}[b]{0.49\textwidth}
    \includegraphics[width=8cm, height=7cm]{fraction_of_opinions_SBM_frac_compl_stubb_REC_8x7.png}
    \caption{SBM}
  \end{subfigure}
  \begin{subfigure}[b]{0.49\textwidth}
    \includegraphics[width=8cm, height=7cm]{fraction_of_opinions_SBM-WS_frac_compl_stubb_REC_8x7.png}
    \caption{SBM-WS}
  \end{subfigure}
  \captionsetup{format=plain}
  \caption[Evolution of the fraction of nodes with opinion 0 ($P_0(t)$) over time for an initial $50/50$ opinion distribution, using the REC filtering algorithm and comparing three different fractions of completely stubborn people (fracRes $= 0$, fracRes $= 0.4$ and fracRes $= 0.8$).]{Evolution of the fraction of nodes with opinion 0 ($P_0(t)$) over time for an initial $50/50$ distribution, using the REC filtering algorithm and comparing three different fractions of completely stubborn people (fracRes $= 0$, fracRes $= 0.4$ and fracRes $= 0.8$). Four different network models are used, each with an average degree around $\left<k\right> \sim 10$ and a total number of nodes $N = 10^3$: \textbf{(a)} Erd\H{o}s-R\'{e}nyi model (ER), $p=0.01$; \textbf{(b)} Watts-Strogatz model (WS) with $K = 10$, $\beta = 0.06$ and $\left<C\right> \sim 0.55$; \textbf{(c)} Stochastic block model (SBM) with 10 communities of size 100, a high modularity case ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001,\ Q \sim 0.9$) and a low modularity case ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q \sim 0.2$) are depicted; \textbf{(d)} Stochastic Watts-Strogatz block model (SBM-WS) with 10 communities of size 100, each community has the following WS parameters: $K = 10,\ \beta = 0.01$ and $p_{\text{add}} = 0.001$, $\left<C\right> \sim 0.55$ and $Q \sim 0.9$. Each plot is the average of $10^2$ independent simulations and data points are shown for every 50 time steps to improve visualization.}%more explanation?
\label{ev_op_50_50_stubb_REC}
\end{figure}

\begin{figure}[H]
  \begin{subfigure}[b]{0.49\textwidth}
  	\includegraphics[width=8cm, height=7cm]{fraction_of_opinions_ER_frac_compl_stubb_PR_8x7.png}
    \caption{ER}
  \end{subfigure}
  \begin{subfigure}[b]{0.49\textwidth}
  	\includegraphics[width=8cm, height=7cm]{fraction_of_opinions_WS_frac_compl_stubb_PR_8x7.png}
    \caption{WS}
  \end{subfigure}
  \begin{subfigure}[b]{0.49\textwidth}
    \includegraphics[width=8cm, height=7cm]{fraction_of_opinions_SBM_frac_compl_stubb_PR_8x7.png}
    \caption{SBM}
  \end{subfigure}
  \begin{subfigure}[b]{0.49\textwidth}
    \includegraphics[width=8cm, height=7cm]{fraction_of_opinions_SBM-WS_frac_compl_stubb_PR_8x7.png}
    \caption{SBM-WS}
  \end{subfigure}
  \captionsetup{format=plain}
  \caption[Evolution of the fraction of nodes with opinion 0 ($P_0(t)$) over time for an initial $50/50$ opinion distribution, using the PR filtering algorithm and comparing three different fractions of completely stubborn people (fracRes $= 0$, fracRes $= 0.4$ and fracRes $= 0.8$).]{Evolution of the fraction of nodes with opinion 0 ($P_0(t)$) over time for an initial $50/50$ distribution, using the PR filtering algorithm and comparing three different fractions of completely stubborn people (fracRes $= 0$, fracRes $= 0.4$ and fracRes $= 0.8$). Four different network models are used, each with an average degree around $\left<k\right> \sim 10$ and a total number of nodes $N = 10^3$: \textbf{(a)} Erd\H{o}s-R\'{e}nyi model (ER), $p=0.01$; \textbf{(b)} Watts-Strogatz model (WS) with $K = 10$, $\beta = 0.06$ and $\left<C\right> \sim 0.55$; \textbf{(c)} Stochastic block model (SBM) with 10 communities of size 100, a high modularity case ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001,\ Q \sim 0.9$) and a low modularity case ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q \sim 0.2$) are depicted; \textbf{(d)} Stochastic Watts-Strogatz block model (SBM-WS) with 10 communities of size 100, each community has the following WS parameters: $K = 10,\ \beta = 0.01$ and $p_{\text{add}} = 0.001$, $\left<C\right> \sim 0.55$ and $Q \sim 0.9$. Each plot is the average of $10^2$ independent simulations and data points are shown for every 50 time steps to improve visualization.}%more explanation?
\label{ev_op_50_50_stubb_PR}
\end{figure}
\noindent
Figure \ref{ev_op_50_50_stubb_PR} shows the results for the evolution of the fraction of opinion 0 in the system for an initial $50/50$ opinion distribution and when using the PR filtering algorithm. Data points are again represented together with the standard deviation $\sigma$. Three different fractions of stubborn people (fracRes $=0$, fracRes $=0.4$ and fracRes $=0.8$) are compared to each other for the four different network models. The results don't differ from those obtained for the REC filtering algorithm. The status-quo is maintained, the initial $50/50$ opinion distribution does not change over time when introducing stubborn people and using the PR filtering algorithm.\\
\newline
Up till now, we have discussed the results for the formation of echo chambers and the evolution of the opinions when introducing a fraction of completely stubborn nodes in the system for an initial $50/50$ opinion distribution. Lets now look at the results for introducing a fraction of stubborn agents when we are dealing with an unbalanced initial opinion distribution, eg. an initial $20/80$ distribution.\\
\newline

@@

>>>>

\subsection{Starting conditions $P_0(0) = 0.5$ and $P_1(0)=0.5$}
\label{50-50}
We start with the following opinion distribution: a fraction of nodes with opinion 0 equal to 0.5 ($P_0(0) = 0.5$) and a fraction of nodes with opinion 1 equal to 0.5 ($P_1(0) = 0.5$). The opinions are randomly distributed over the network. The opinion dynamics model discussed in Chapter \ref{chap4}, Section \ref{opinionmodel} % correct section??
is used. We set the activation probability equal to $p_{\text{act}} = 0.1$ and the actual time-line of each user has a maximum length equal to $S=20$.\\
\newline
Four different network models are implemented, each with a total number of nodes equal to $N = 10^3$ and an average degree around $\left<k\right> \sim 10$
\begin{itemize}
	\item The Erd\H{o}s-R\'{e}nyi model (ER) with edge probability $p = 0.01$.
	\item The Watts-Strogatz model (WS) with mean degree $K=10$; two different values of the rewire probability $\beta$ are used, each of them has a slightly different value of the average clustering coefficient, namely $\beta = 0.01$ with $\left<C\right> \sim 0.60$ and $\beta = 0.06$ with $\left<C\right> \sim 0.55$. This model is used to investigate the effect of clustering.
	\item The stochastic block model (SBM) with 10 communities of size 100; two networks with different modularities $Q$ are investigated: a high modularity network, $Q \sim 0.9$, with an edge probability inside the communities equal to $p_{\text{cl}} = 0.1$ and an edge probability between communities equal to $p_{\text{add}} = 0.001$, and a low modularity network ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q \sim 0.2$). This model is used to investigate the effect of a community structure.
	\item The stochastic Watts-Strogatz block model (SBM-WS) with 10 communities of size 100. Each community has the following Watts-Strogatz parameters: mean degree $K =10$ and rewire probability $\beta = 0.01$. Edges are added between communities with a probability $p_{\text{add}} = 0.001$. This is a network with both a high clustering ($\left<C\right> \sim 0.55$) and a high modularity ($Q \sim 0.9$).
\end{itemize}
\textbf{Evolution of the opinion fractions}\\
\newline
We start with the results for the evolution of the opinion fractions in case of an initial $50/50$ opinion distribution. Prof. N. Perra and prof. Luis E.C. Rocha already showed that, for the case of balanced initial opinions, the prevalence of the opinions do not change over time for neither the random network, nor the Watts-Strogatz network and for none of the different filtering algorithms \cite{Perra2019}. In Figure \ref{ev_op_50_50} the results for the network model with a community structure (SBM, Figure \ref{sbm_50-50}) and the network model with both a community structure and a high clustering (SBM-WS, Figure \ref{sbm-ws_50-50}) are shown for both the REC and the PR filtering algorithm. Results are shown together with their standard deviation $\sigma$. From this figure it becomes clear that having a community structure is also not able to break the status-quo for both of the filtering algorithms.

\begin{figure}[H]
  \begin{subfigure}[b]{0.49\textwidth}
    \includegraphics[width=8cm, height=7cm]{fraction_of_opinions_SBM_10x100_01-0001_003-0008_50-50_PR_REC.png}
    \caption{SBM}
    \label{sbm_50-50}
  \end{subfigure}
  \begin{subfigure}[b]{0.49\textwidth}
    \includegraphics[width=8cm, height=7cm]{fraction_of_opinions_SBM-WS_10x100_10-001-0001_50-50_PR_REC.png}
    \caption{SBM-WS}
    \label{sbm-ws_50-50}
  \end{subfigure}
  \captionsetup{format=plain}
  \caption[Evolution of the fraction of nodes with opinion 0 ($P_0(t)$) over time for an initial $50/50$ opinion distribution.]{Evolution of the fraction of nodes with opinion 0 ($P_0(t)$) over time for an initial $50/50$ distribution. Two different network models with a community structure are represented, each with an average degree around $\left<k\right> \sim 10$ and a total number of nodes $N = 10^3$: \textbf{(a)} Stochastic block model (SBM) with 10 communities of size 100, a high modularity case ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001,\ Q \sim 0.9$) and a low modularity case ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q \sim 0.2$) are depicted; \textbf{(b)} Stochastic Watts-Strogatz block model (SBM-WS) with 10 communities of size 100, each community has the following WS parameters: $K = 10,\ \beta = 0.01$ and $p_{\text{add}} = 0.001$, $\left<C\right> \sim 0.55$ and $Q \sim 0.9$. For each model two different filtering algorithms are compared to each other: the REC and the PR filtering algorithm. Each plot is the average of $10^2$ independent simulations and data points are shown for every 50 time steps to improve visualization.}
\label{ev_op_50_50}
\end{figure}
\noindent
\textbf{Echo chambers}\\
\newline
Figure \ref{echo_50_50} shows the distribution of the fraction of friends (nearest neighbors, nn) of node $i$ with the same opinion 1 at $t=500$, $\left<P_1^{\text{nn}}\right>$. The value shown on the $y$-axis, $F_N(\left<P_1^{\text{nn}}\right>)$, is the result of dividing the average distribution of opinions in each neighborhood at $t=500$ by its corresponding value at $t=0$. The region on the $x$-axis close to zero describes neighborhoods in which none, or very few users, adopt opinion 1 (these are thus neighborhoods in which most users adopt opinion 0). Conversely, the region on the $x$-axis close to one describes neighborhoods in which a majority of the users adopt opinion 1. The distribution is obtained by imposing ten equispaced bins, $\Delta x = 0.1$ \cite{Perra2019}. Error bars are omitted for the sake of an improved visualization.\\
These types of plots are perfectly suited to investigate the formation of opinion clusters (echo chambers) and effects such as polarization in the system under study.\\

@@

@@

\begin{figure}[H]
  \begin{subfigure}[b]{0.49\textwidth}
  	\includegraphics[width=8cm, height=7cm]{fraction_of_opinions_ER_001_50-50_PR_REC_20-80_8x7.png}
    \caption{ER}
  \end{subfigure}
  \begin{subfigure}[b]{0.49\textwidth}
  	\includegraphics[width=8cm, height=7cm]{fraction_of_opinions_WS_10-006_PR_REC_20-80_8x7.png}
    \caption{WS}
    \label{ws_20-80_op}
  \end{subfigure}
  \begin{subfigure}[b]{0.49\textwidth}
    \includegraphics[width=8cm, height=7cm]{fraction_of_opinions_SBM_01-0001_003-0008_PR_REC_20-80_8x7.png}
    \caption{SBM}
  \end{subfigure}
  \begin{subfigure}[b]{0.49\textwidth}
    \includegraphics[width=8cm, height=7cm]{fraction_of_opinions_SBM-WS_10-001-0001_PR_REC_20-80_8x7.png}
    \caption{SBM-WS}
    \label{sbm-ws_20-80_op}
  \end{subfigure}
  \captionsetup{format=plain}
  \caption[Evolution of the fraction of nodes with opinion 0 ($P_0(t)$) over time for an initial $20/80$ opinion distribution.]{Evolution of the fraction of nodes with opinion 0 ($P_0(t)$) over time for an initial $20/80$ distribution. Four different network models are used, each with an average degree around $\left<k\right> \sim 10$ and a total number of nodes $N = 10^3$: \textbf{(a)} Erd\H{o}s-R\'{e}nyi model (ER), $p=0.01$; \textbf{(b)} Watts-Strogatz model (WS) with $K = 10$, $\beta = 0.06$ and $\left<C\right> \sim 0.55$; \textbf{(c)} Stochastic block model (SBM) with 10 communities of size 100, a high modularity case ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001,\ Q \sim 0.9$) and a low modularity case ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q \sim 0.2$) are depicted; \textbf{(d)} Stochastic Watts-Strogatz block model (SBM-WS) with 10 communities of size 100, each community has the following WS parameters: $K = 10,\ \beta = 0.01$ and $p_{\text{add}} = 0.001$, $\left<C\right> \sim 0.55$ and $Q \sim 0.9$. For each model two different filtering algorithms are compared to each other: the REC and the PR filtering algorithm. Each plot is the average of $10^2$ independent simulations and data points are shown for every 50 time steps to improve visualization.}
\label{ev_op_20_80}
\end{figure}

@@

  	\includegraphics[width=8cm, height=7cm]{fraction_of_opinions_WS_1000-6_PR_20-80_8x7.png}
    \caption{WS, $N = 10^3,\ K=6,\ \beta=0.01$}
    \label{ws1000}

@@

  	\includegraphics[width=8cm, height=7cm]{fraction_of_opinions_WS_10000-6_PR_20-80_8x7.png}
    \caption{WS, $N = 10^4,\ K=6,\ \beta=0.01$}
    \label{ws10000}

@@

  \caption[Evolution of the fraction of nodes with opinion 0 ($P_0(t)$) over time for an initial $20/80$ opinion distribution for a WS model with $N=10^3$ or $N=10^4$, $K=6$ and $\beta = 0.01$, PR.]{Evolution of the fraction of nodes with opinion 0 ($P_0(t)$) over time for an initial $20/80$ distribution. The Watts-Strogatz network is depicted with \textbf{(a)} $N=10^3,\ K=6,\ \beta=0.01$ and \textbf{(b)} $N=10^4,\ K=6,\ \beta=0.01$. In both cases the average clustering coefficient is around $\left<C\right> \sim 0.58$ and the PR filtering method is used. Each plot is the average of $10^2$ independent simulations and data points are shown for every 50 time steps to improve visualization.}
\label{ev_op_WS_k=6}

@@

@@

Lets first discuss the results for the Erd\H{o}s-R\'{e}nyi model and the stochastic block model, see Figure \ref{er_echo_20-80} and Figure \ref{sbm}. When using the REC method, the formation of echo chambers around the majority opinion 1 is hampered. This is somewhat surprising, since $80\%$ of the nodes carry this opinion and is probably a consequence of the absence of clustering and the presence of short average path lengths in these models. There is even a little spike around $\left<P_1^{\text{nn}}\right> = 0.1$, which indicates a small increase in the fraction of nodes with the majority of their neighbors carrying the minority opinion 0. From figure \ref{ev_op_20_80} it was already concluded that the REC method doesn't break the prevalence of the opinions. The opinion distribution is however slightly shifted from the case where the minority opinion 0 is more or less randomly distributed over the network to the case where the minority opinion 0 groups together in opinion clusters. \\

@@

\begin{figure}[H]
  \begin{subfigure}[b]{0.49\textwidth}
  	\includegraphics[width=8cm, height=7cm]{echo_chambers_no_stubb_ER_001_20-80_PR_REC.png}
    \caption{ER}
    \label{er_echo_20-80}
  \end{subfigure}
  \begin{subfigure}[b]{0.49\textwidth}
  	\includegraphics[width=8cm, height=7cm]{echo_chambers_no_stubb_WS_10_006_20-80_PR_REC_8x7.png}
    \caption{WS}
    \label{ws}
  \end{subfigure}
  \begin{subfigure}[b]{0.49\textwidth}
    \includegraphics[width=8cm, height=7cm]{echo_chambers_no_stubb_SBM_01-0001_003-0008_10x100_20-80_PR_REC_8x7.png}
    \caption{SBM}
  	\label{sbm}    
  \end{subfigure}
  \begin{subfigure}[b]{0.49\textwidth}
    \includegraphics[width=8cm, height=7cm]{echo_chambers_no_stubb_SBM-WS_10_001_0001_10x100_20-80_PR_REC_8x7.png}
    \caption{SBM-WS}
    \label{sbm-ws}
  \end{subfigure}
  \captionsetup{format=plain}
  \caption[Distribution of the fraction of friends (nearest neighbors, nn) of node $i$ with the same opinion 1 at $t=500$, $\left<P_1^{\text{nn}}\right>$, for an initial $20/80$ opinion distribution.]{Distribution of the fraction of friends (nearest neighbors, nn) of node $i$ with the same opinion 1 at $t = 500$. The distribution is normalized by dividing for the same quantity computed at $t=0$. Four different network models are used, each with a total number of nodes $N = 10^3$ and an average degree around $\left<k\right> \sim 10$: \textbf{(a)} Erd\H{o}s-R\'{e}nyi model (ER), $p=0.01$; \textbf{(b)} Watts-Strogatz model (WS) with $K = 10$, $\beta = 0.06$ and $\left<C\right> \sim 0.55$; \textbf{(c)} Stochastic block model (SBM) with 10 communities of size 100, a high modularity case ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001,\ Q \sim 0.9$) and a low modularity case ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q \sim 0.2$) are depicted; \textbf{(d)} Stochastic Watts-Strogatz block model (SBM-WS) with 10 communities of size 100, each community has the following WS parameters: $K = 10,\ \beta = 0.01$ and $p_{\text{add}} = 0.001$, $\left<C\right> \sim 0.55$ and $Q \sim 0.9$. For each model two different filtering algorithms are compared to each other: the REC and the PR filtering algorithm. The starting conditions are $P_0(0) = 0.2$ and $P_1(0) = 0.8$. Each plot is the average of $10^2$ independent simulations.}
\label{echo_20_80}
\end{figure}

% talk about small spike around 1 for PR method??
% explain black dotted line??

@@

The short answer is no. Introducing stubborn people in the system does not affect the evolution of the opinions over time for neither the REC, nor the PR filtering algorithm. The balanced opinion distribution is maintained over time. This result is valid for all the four implemented network models (ER, WS, SBM and SBM-WS). Network topologies such as a high clustering coefficient and/or a community structure do not change this result. \\% Include figure??

@@

Up till now, we have discussed the results for the formation of echo chambers and the evolution of the opinions when introducing a fraction of completely stubborn nodes in the system for an initial $50/50$ opinion distribution. Lets now look at the results for introducing a fraction of stubborn agents when we are dealing with an unbalanced initial opinion distribution, eg. an initial $20/80$ distribution.\\

@@

  	\includegraphics[width=8cm, height=7cm]{fraction_of_opinions_ER_frac_compl_stubb_PR_20-80_8x7.png}
  	\includegraphics[width=8cm, height=7cm]{fraction_of_opinions_WS_frac_compl_stubb_PR_20-80_8x7.png}
    \includegraphics[width=8cm, height=7cm]{fraction_of_opinions_SBM_frac_compl_stubb_PR_20-80_8x7.png}
    \includegraphics[width=8cm, height=7cm]{fraction_of_opinions_SBM-WS_frac_compl_stubb_PR_20-80_8x7.png}
  \caption[Evolution of the fraction of nodes with opinion 0 ($P_0(t)$) over time for an initial $20/80$ opinion distribution, using the PR filtering algorithm and comparing three different fractions of completely stubborn people (fracRes $= 0$, fracRes $= 0.4$ and fracRes $= 0.8$).]{Evolution of the fraction of nodes with opinion 0 ($P_0(t)$) over time for an initial $20/80$ distribution, using the PR filtering algorithm and comparing three different fractions of completely stubborn people (fracRes $= 0$, fracRes $= 0.4$ and fracRes $= 0.8$). Four different network models are used, each with an average degree around $\left<k\right> \sim 10$ and a total number of nodes $N = 10^3$: \textbf{(a)} Erd\H{o}s-R\'{e}nyi model (ER), $p=0.01$; \textbf{(b)} Watts-Strogatz model (WS) with $K = 10$, $\beta = 0.06$ and $\left<C\right> \sim 0.55$; \textbf{(c)} Stochastic block model (SBM) with 10 communities of size 100, a high modularity case ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001,\ Q \sim 0.9$) and a low modularity case ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q \sim 0.2$) are depicted; \textbf{(d)} Stochastic Watts-Strogatz block model (SBM-WS) with 10 communities of size 100, each community has the following WS parameters: $K = 10,\ \beta = 0.01$ and $p_{\text{add}} = 0.001$, $\left<C\right> \sim 0.55$ and $Q \sim 0.9$. Each plot is the average of $10^2$ independent simulations and data points are shown for every 50 time steps to improve visualization.}%more explanation?
\label{ev_op_20_80_stubb_PR}
As for the balanced initial conditions, introducing stubborn people does not break the prevalence of the initial opinions for the REC filtering algorithm. Introducing stubborn people in the system combined with the PR filtering algorithm, on the other hand, does have an impact in the case of unbalanced initial conditions.\\
